Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2011-12-04 ltx/paper.tex
been working on draft of paper... better intro, edits throughout

\documentclass[smallcondensed, final]{svjour3}

@@

We can define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: differentiating the spatiotemporal regions of a video that constitute objects from the region that constitutes the background (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the positions of the detected objects over time (`tracking'). An ability to carry out these tasks in an automated manner is useful for many disciplines that make use of video data, including the fields of surveillance, autonomous systems, robotics, computational biology, and video summarization. Algorithms of this sort that prove to be efficacious for many videos and a variety of objects are particularly desirable. This paper serves to introduce a new technique involving the application of a dependent Dirichlet processes mixture model, which we hope will provide the foundation for a number of robust, unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research over the past ten years related to the development of methods for general detection and tracking of objects has tended to focus on one of extraction, localization, or tracking; attempts to integrate in the other two and produce a full system able to handle arbitrary objects and videos was often not a primary focus, and was often attempted afterwards, in an ad-hoc manner. Research aimed at developing methods to accomplish the three tasks in a cohesive manner have been pursued more recently to produce algorithms that can, in a fully automated way, accomplish the goal of deciding which spatiotemporal regions of a video constitute distinct objects. A widely applied, though unsophisticated, example is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold to extract pixel locations, and segmenting the extracted data into `blobs' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting occluded objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.

%Note that blob tracking carries out the three main tasks described above, so it satisfies the criteria for the type of unsupervised detection and tracking algorithms studied in this work;

This paper presents the use of Dirichlet process mixture models as a technique for unsupervised multiple object detection and tracking, which we intend as a method to allow for robust object detection (particularly in cases where frame-by-frame segmentation would be difficult) and to keep the tracks of independent objects isolated. Specifically, this work applies a type of Bayesian nonparametric model to data gained during extraction; note that data of this type is refered to in the paper as `extraction data', `data points', or `observations'. The model is first presented in an abstract form, which could be adapted to a wide range of extraction procedures and object appearance representations (mainly left for exploration in future studies). We then describe inference algorithms that allow object localization and tracking to be carried out. Afterwards, we formulate a definite model for data gained during a specific extraction procedure and implement the described inference procedures to carry out multiple object detection and tracking in a number of experiments.

\subsection{Related Work}

Related work can be broken down into studies relating to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.

The task of extraction incorporates the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes. These procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. These pixel modeling techniques have been adapted to allow for time-dependent probability distributions over pixel values, which help account for slighlty changing backgrounds, such as those caused from brief movements of background objects (for example, trees in outdoor scenes). Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Post processing has been applied to add information regarding the edges of objects in an attempt to decrease fragmentation \citep{turdu_2007}. Others have attempted to achieve better foreground modeling in cases where the foreground and background exhibit similar color distributions (the so called `color similarity problem') by, for example, incorporating models of the foreground into a typical background model (such as the one described above) \citep{gallego_2009, mchugh_2009}.

On the other hand, heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. While these methods locate regions with objects less accurately than model-based methods, they are often much less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition to the two techniques mentioned above, there exists other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

The task of localization is closely related to the topics of object detection, position-finding, target segmentation, contour-detection, template matching, and recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

The task of tracking incorporates the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain localization of the positions and/or shapes of objects over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object. A great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.

There are an extensive collection of methods for tracking single, non-rigid, objects. For example, color has long been used for tracking. In this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified. In general, these schemes operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}. Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.

@@

\section{Model Overview}
\label{sec:modeloverview}
%
This section provides details on the nature of the data to be modeled, Dirichlet process mixture (DPM) models, and dependent Dirichlet process mixture (DDPM) models. We end by making explicit the way in which characteristics of video objects can be represented by elements of these models. We formulate the model in a sufficently abstract manner to allow for specific implementations tailored towards specified objects in future studies.

\subsection{Data}
\label{sec:data}
%
We defined the task of extraction to be the unsupervised decision of which spatiotemporal regions of a video constitute objects; hence, the extraction data modeled in this work are D-dimensional observations that in some way provide information about the regions of the video frame that contain objects. A simple example of this information is shown by the three-dimensional extraction data $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of pixels contained in regions comprising objects in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and should always be kept as features of the data for modeling (as position information is key for reasonable tracking and time information is necessary for our intention to model the temporal dependencies of the data).

Examples of additional features that could potentially be extracted from video frame regions containing objects include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform results (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and, when viewed in cumulation for multiple observations, are distributed in a way that allows for modeling and tractable inference (this is discussed further in section 3).

We will use the phrase `baseline' three-dimensional extraction data to refer to the $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). An example of data with such a distribution can been seen in Figure \ref{fig:baselinedata}. The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
\begin{figure*}[h]
        \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
        \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants. The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$ spatial coordinates, respectively, of pixels within a given frame. Markers are colored by position for ease of viewing.}
\end{figure*}
Inference performed on the time dependent Dirichlet process mixture model (described in the following sections) carries out clustering of the data, which allows each worm-like data cloud to be isolated and independent objects to be localized and tracked.

@@

Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models. These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis. They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models. In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video. Consequentially, a model which is able to perform robust nonparametric estimation of the number of cluster is well suited to be applied to the task of unsupervised object detection and tracking.

The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup. We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, presently, which allows us to subsequently describe DPMs in context.

@@

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (where the coefficients of the linear combination sum to one); each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form. The finite mixture model can be written as

\begin{equation}
P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
\end{equation}

where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$ (note that these coefficients now sum to one naturally due to the nature of probability distributions). Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component. We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $. This model can also be written hierarchically as
\begin{align}
\begin{split}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$).

@@

The finite mixture model of the previous section can be extended to a Bayesian perspective by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights. The resulting Bayesian mixture model can be formulated in a hierarchical representation as
\begin{align}
\begin{split}
	\label{bayesian_mixture_model}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from their prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from their prior distribution, where `Dir' refers to a standard Dirichlet distribution.

@@

The Dirichlet process (DP), first introduced by \cite{ferguson_1973} in 1973, may be intuitively viewed as a probability distribution over discrete probability distributions. Accordingly, draws from a DP are probability mass functions (PMFs). A DP is parameterized by a base distribution $\mathbb{G}_{0}$, which is a probability distribution over a set $\Theta$, and a concentration parameter $\alpha \in \mathbb{R}_{+}$. We say that $G$ is a random PMF distributed according to a DP, written $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, if the following holds for all finite partitions $A_{1}, \ldots, A_{p}$ of $\Theta$:
\begin{equation}
(G(A_{1}), \ldots, G(A_{p})) \sim \text{Dir}(\alpha \mathbb{G}_{0}(A_{1}), \ldots, \alpha \mathbb{G}_{0}(A_{p}))
\end{equation}
Where `Dir' denotes a standard Dirichlet distribution. The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP. This is due to the fact that if the base distribution $\mathbb{G}_{0}$ is a distribution over $\Theta$, $A \subset \Theta$, and $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, then the following holds:
\begin{equation}
\mathbb{E}[G(A)] = \mathbb{G}_{0}(A)
\end{equation}
\begin{equation}
\text{Var}[G(A)] = \mathbb{G}_{0}(A) (1 - \mathbb{G}_{0}(A)) / (\alpha + 1)
\end{equation}
Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.

@@

To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at adjacent positions in the sequence. For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps. Research towards models which accomplish such goals are described in (list citations).

>>>>

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[twocolumn, final]{svjour3}

@@

We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking'). An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, computational biology, and video summarization. Algorithms proving to be efficacious for a variety of objects and video types are particularly applicable. This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking--or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into `blobs' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking}, preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for unsupervised multiple object detection and tracking. This method is intended to allow for robust detection of arbitrary objects (particularly in cases where frame-by-frame segmentation is difficult) and better isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how this generative model may be applied to types of noisy data that can be quickly extracted from objects in videos, and how this model allows for automated detection and tracking. The model is first presented in an abstract form, to allow for it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.

\subsection{Related Work}
\label{sec:relatedwork}

Related work involves studies related to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.

The task of extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes. These procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. These methods are often less accurate than model-based methods, but also far less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition to the two techniques mentioned above, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

The task of localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering \willie{must cite here}. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

The task of tracking incorporates the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.

Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking. These techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times \willie{cite paper(s) for each of these}.

In the past five years, there has been a push for methods that incorporate detection and tracking of multiple objects to allow for full systems capable of autonomously handling real-world video datasets.
\willie{Describe tracking by detection, more on blob tracking, the other few methods i've found for automated joint detection and tracking}

% Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.

@@

\section{Model Overview}
\label{sec:modeloverview}
%
This section provides details on the data to be modeled, the Dirichlet process mixture (DPM) model, and the dependent Dirichlet process mixture (DDPM) model. We make explicit the way in which characteristics of video objects can be represented by elements of these models. We formulate the model in a sufficently abstract manner to allow for implementations tailored for specified objects in future studies.

\subsection{Data}
\label{sec:data}
%
We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}. The features of the data are characteristics of pixels in the regions of the video that contain objects, and the data are gained during extraction. For example, at its most basic, an extraction dataset is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region representing an object in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and are necessary features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).

Examples of additional features that could potentially be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference_spec}).

We will use the phrase `baseline' three-dimensional extraction data to refer to the $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}. The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
\begin{figure*}[h]
        \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
        \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants. The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$, the spatial coordinates of pixels within a given frame. Markers are colored by position for ease of viewing.}
\end{figure*}
Inference performed on the time dependent Dirichlet process mixture model (described in Section~\ref{sec:inference_spec}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.

@@

Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models. These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis. They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models. In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video. Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in an extraction dataset may be applied to perform unsupervised object detection and tracking.

The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup. We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.

@@

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (where the coefficients of the linear combination sum to one); each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form. The finite mixture model can be written as
\begin{equation}
P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
\end{equation}
where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$. Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component. We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $. This model can also be written hierarchically as
\begin{align}
\begin{split}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$).

@@

The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights. The resulting Bayesian mixture model can be formulated in a hierarchical representation as
\begin{align}
\begin{split}
	\label{bayesian_mixture_model}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from the prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.

@@

The Dirichlet process (DP), first introduced by \cite{ferguson_1973} in 1973, may be intuitively viewed as a probability distribution over discrete probability distributions. Accordingly, draws from a DP are probability mass functions (PMFs). A DP is parameterized by a base distribution $\mathbb{G}_{0}$, which is a probability distribution over a set $\Theta$, and a concentration parameter $\alpha \in \mathbb{R}_{+}$. We say that $G$ is a random PMF distributed according to a DP, written $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, if the following holds for all finite partitions $A_{1}, \ldots, A_{p}$ of $\Theta$:
\begin{equation}
(G(A_{1}), \ldots, G(A_{p})) \sim \text{Dir}(\alpha \mathbb{G}_{0}(A_{1}), \ldots, \alpha \mathbb{G}_{0}(A_{p}))
\end{equation}
Where `Dir' denotes a Dirichlet distribution. The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP. This is due to the fact that if the base distribution $\mathbb{G}_{0}$ is a distribution over $\Theta$, $A \subset \Theta$, and $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, then the following holds:
\begin{equation}
\mathbb{E}[G(A)] = \mathbb{G}_{0}(A)
\end{equation}
\begin{equation}
\text{Var}[G(A)] = \mathbb{G}_{0}(A) (1 - \mathbb{G}_{0}(A)) / (\alpha + 1)
\end{equation}
Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.

@@

To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at neighboring positions in the sequence. For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps \willie{cite some of these}.
